---
title: "MachineLearning_Project"
author: "Ronald Armando"
date: "4/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Loading and Exploratory Analysis

## Dataset Overview
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project can be found here: http://groupware.les.inf.puc-rio.br/har.

A short description of the datasets content from the authors’ website:

“Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg)."

## Getting and loading the data
We first upload the R libraries that are necessary for the complete analysis.

```{r}
rm(list=ls())               
library(knitr)
library(caret)
library(randomForest)
library(corrplot)
library(rpart)
library(rpart.plot)
library(rattle)


set.seed(12345)

train_Url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_Url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(train_Url), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(test_Url), na.strings=c("NA","#DIV/0!",""))
```

